Compute Servers for Teaching Big Data

### 2025-01-28 paper.tex
section 1

\section{Introduction}

@@

\subsection{The Course}

We developed and have taught the course Stat 129, titled `Analyzing and Processing Big Data'.
This is an upper division elective course offered by the Mathematics and Statistics Department, and is required for students pursuing a bachelor's degree in Statistics with an emphasis in Data Science.
The catalog course description is:

@@

Students typically take this course their junior year after taking Stat 128, Statistical Computing, an upper division requirement for the statistics major focused on the R programming language, with content similar to Data Science in a Box, R4DS TODO: Cite.

In Stat 129 we use and teach technologies including bash, Python, and SQL.
Our intention is to expose students broadly to the common tools in data science.
Our general approach to data science education is to provide students with real world data sets and ask open ended questions, squarely in the tradition of Deborah Nolan and Duncan Temple Lang \cite{nolan2010computing}.
Indeed, they were our teachers and mentors.

@@

It's essential that students access data and run code on remote servers, because there is a qualitative difference between running code on a laptop that's sitting in front of you and a machine that you never actually see, that you can only access through ssh, secure shell.
We teach bash, the classic command line, because it has proven to be a durable technology; it's been around since the 1970's, and it isn't going away any time soon.
In general, students benefit when instructors focus on proven core technologies, rather than on what happens to be popular this year, because they come away with skills that they can apply in many different contexts.

>>>>

@@

\section{The Course}

Stat 129, `Analyzing and Processing Big Data', is an upper division elective course offered by the Mathematics and Statistics Department at CSU Sacramento, and is required for students pursuing a bachelor's degree in Statistics with an Emphasis in Data Science.
The catalog course description is:

@@

Stat 129 uses technologies including bash, Python, and SQL to analyze large data sets, with the intention of exposing students to the common tools in data science.
Students tend to be highly motivated, because many of them hope to land jobs requiring these skills after graduation.
Stat 129 requires a programming prerequisite, since it is too advanced for a first programming course.
Statistics majors take Stat 129 after taking `Statistical Computing', an upper division course focused on data analysis using the R programming language, with content similar to Data Science in a Box, R4DS TODO: Cite.
%Ideally, students would have taken two semesters of programming before this course.

The philosophy behind this class is to provide students with real world data sets and ask open ended questions, following in the footsteps of our teachers and mentors, Deborah Nolan and Duncan Temple Lang \cite{nolan2010computing}.
We have been teaching data science programming technologies in an academic setting since 2015, and we developed Stat 129 and have taught it each year since 2021.
Feedback from this data forward approach in general, and this class in particular, has been overwhelmingly positive.
In Stat 129 students ask real questions based on real data that resides on a real server, and for this reason several alumni have reported that this has been the single most practical and valuable class that they have taken in college.

@@

It's essential that students access data and run code on remote servers for a class on `Big Data', because there is a qualitative difference between running code on a laptop that's sitting in front of you and a machine that you never actually see, that you can only access through ssh, secure shell.
We teach bash, the classic command line, because it has proven to be a durable technology; it's been around since the 1970's, and it isn't going away any time soon.
In general, students benefit when instructors focus on proven core technologies, rather than on what happens to be popular this year, because they come away with skills that they can apply in many different contexts.
